---
title: "vignette atime"
author: "Doris Amoakohene"
date: "`r Sys.Date()`"
output: html_document
---


Modify atime compare-data.table-tidyverse vignette, and analyze efficiency of packages such as polars, arrow, collapse, spark.
To modify the atime vignette in the compare-data.table-tidyverse package and analyze the efficiency of packages such as polars, arrow, collapse, and spark, you would need to perform the following steps:

Perform Comparison with 
data.table
arrow

adding:

collapse 
spark


The purpose of this vignette is to make figures which show the efficiency of data.table.

fwrite: Fast csv writer

```{r}
library(data.table)
library(readr)
library(arrow)
library(ggplot2)
library(collapse)
#library(sparklyr)
library(polars)
library(dplyr)
library(plyr)
library(tidyr)
library(stats)

#sc <- sparklyr::spark_connect(master = "local")
```

```{r}
write.colors <- c(
  "readr::write_csv"="#9970AB",
  "data.table::fwrite"="#D6604D",
  "write_csv_arrow"="#BF812D", 
  "polars::write_csv"="#33A02C",
  "write_CSV_COllapse" = "#722f37",
  #"write_csv_spark"= "pink",
  "write.csv2"= "#1F78B4",
  "utils::write.csv"="deepskyblue")

n.rows <- 100
seconds.limit <- 1


atime.write.vary.cols <- atime::atime(
  N=as.integer(10^seq(2, 6, by=0.5)),
  setup={
    set.seed(1)
    input.vec <- rnorm(n.rows*N)
    input.mat <- matrix(input.vec, n.rows, N)
    input.df <- data.frame(input.mat)
    #spark_df<- copy_to(sc,input.mat, name = "spark_df")
    
  },
  seconds.limit = seconds.limit,
  "data.table::fwrite"={
    data.table::fwrite(input.df, tempfile(), showProgress = FALSE)
  },
  "write_csv_arrow"={
    arrow::write_csv_arrow(input.df, tempfile())
  },
  "readr::write_csv"={
    readr::write_csv(input.df, tempfile(), progress = FALSE)
  },
  "polars::write_csv" = {
    write_csv(input.df, tempfile())
  },
  "write_csv_collapse"={
    write.csv(input.df,tempfile())
  },
  #"write_csv_spark"={
    #spark_write_csv(spark_df, tempfile(), mode = "overwrite")
  #},
  "write.csv2" = {
    write.csv2(input.df, tempfile())
  },
  "utils::write.csv"= {
    utils::write.csv(input.df, tempfile())
  }
)
```



```{r}
refs.write.vary.cols <- atime::references_best(atime.write.vary.cols)
pred.write.vary.cols <- predict(refs.write.vary.cols)

gg.write <- plot(pred.write.vary.cols)+
  theme(text=element_text(size=20))+
  ggtitle(sprintf("Write real numbers to CSV, %d x N", n.rows))+
  scale_x_log10("N = number of columns to write")+
  scale_y_log10("Computation time (seconds)
median line, min/max band
over 10 timings")+
  facet_null()+
  scale_fill_manual(values=write.colors)+
  scale_color_manual(values=write.colors)

```

```{r}
gg.write
```

fread: fast CSV reader

```{r}
read.colors <- c(
  "readr::read_csv\n(lazy=TRUE)"="#9970AB",
  "readr::read_csv\n(lazy=FALSE)"="#9970AB",
  "data.table::fread"="#D6604D",
  "read_csv_arrow"="#BF812D",
  "polars::read_csv"="#33A02C",
  "read_csv_collapse"="#722f37",
  "read.csv2" = "#1F78B4",
  "utils::read.csv"="deepskyblue")

atime.read.vary.cols <- atime::atime(
  N=as.integer(10^seq(2, 6, by=0.5)),
  setup={
    set.seed(1)
    input.vec <- rnorm(n.rows*N)
    input.mat <- matrix(input.vec, n.rows, N)
    input.df <- data.frame(input.mat)
    input.csv <- tempfile()
    fwrite(input.df, input.csv)
  },
  seconds.limit = seconds.limit,
  "data.table::fread"={
    data.table::fread(input.csv, showProgress = FALSE)
  },
  "read_csv_arrow"={
    arrow::read_csv_arrow(input.csv)
  },
  "readr::read_csv\n(lazy=TRUE)"={
    readr::read_csv(input.csv, progress = FALSE, show_col_types = FALSE, lazy=TRUE)
  },
  "readr::read_csv\n(lazy=FALSE)"={
    readr::read_csv(input.csv, progress = FALSE, show_col_types = FALSE, lazy=FALSE)
  },
  "polars::read_csv" = {
    read_csv(input.csv)
  },
  "read_csv_collapse"={
    read.csv(input.csv)
  },
   "read.csv2" = {
    read.csv2(input.csv)
   },
  "utils::read.csv"=utils::read.csv(input.csv))
```


```{r}
refs.read.vary.cols <- atime::references_best(atime.read.vary.cols)
pred.read.vary.cols <- predict(refs.read.vary.cols)

gg.read <- plot(pred.read.vary.cols)+
  theme(text=element_text(size=20))+
  ggtitle(sprintf("Read real numbers from CSV, %d x N", n.rows))+
  scale_x_log10("N = number of columns to read")+
  scale_y_log10("Computation time (seconds)
median line, min/max band
over 10 timings")+
  facet_null()+
  scale_fill_manual(values=read.colors)+
  scale_color_manual(values=read.colors)
```

```{r}
gg.read
```


Summarize by group



```{r}
ml.colors <- c(
  "dplyr::summarise"="#9970AB",
  "[.data.table"="#D6604D",
  "stats::aggregate"="deepskyblue",
  "plyr::ddply"="orange",
  "tidyr::pivot_longer"="green")
options(dplyr.summarise.inform=FALSE)
n.folds <- 10
ml.atime <- atime::atime(
  N=as.integer(10^seq(2, 7, by=0.5)),
  setup={
    loss.dt <- data.table(
      name="loss", 
      fold=rep(1:n.folds, each=2*N),
      loss=rnorm(2*N*n.folds),
      set=rep(c("subtrain","validation"),each=N),
      epoch=1:N,
      key=c("set","epoch","fold"))
  },
  seconds.limit=seconds.limit,
  "[.data.table"={
    loss.dt[, .(
      loss_length=.N,
      loss_mean=mean(loss),
      loss_sd=sd(loss)
    ), by=.(set, epoch)]
  },
  "stats::aggregate"={
    res <- stats::aggregate(
      loss ~ set + epoch, 
      loss.dt, 
      function(values)list(c(
        loss_length=length(values),
        loss_mean=mean(values), 
        loss_sd=sd(values))))
    data.frame(
      subset(res, select=-loss), 
      do.call(rbind, res$loss))
  },
  "plyr::ddply"={
    ddply(loss.dt, c("set", "epoch"), summarize,
          loss_length = length(loss),
          loss_mean = mean(loss),
          loss_sd = sd(loss))
  },
  "tidyr::pivot_longer"={
    tidy_data <- pivot_longer(loss.dt, cols = starts_with("loss"), names_to = "GroupVar", values_to = "NumericVar")
    
    summary_data <- tidy_data %>%
      group_by(set, epoch) %>%
      summarise(
        loss_length = length(NumericVar),
        loss_mean = mean(NumericVar),
        loss_sd = sd(NumericVar)
      )
    },
  "dplyr::summarise"={
    loss.dt |> 
      dplyr::group_by(set, epoch) |> 
      dplyr::summarise(
        loss_length=length(loss),
        loss_mean=mean(loss), 
        loss_sd=sd(loss))
  })

```


```{r}
ml.refs <- atime::references_best(ml.atime)
ml.pred <- predict(ml.refs)
ml.gg <- plot(ml.pred)+
  theme(text=element_text(size=20))+
  ggtitle(sprintf("Mean,SD,Length over %d real numbers, N times", n.folds))+
  scale_x_log10("N = number of Mean,SD,Length to compute")+
  scale_y_log10("Computation time (seconds)
median line, min/max band
over 10 timings")+
  facet_null()+
  scale_fill_manual(values=ml.colors)+
  scale_color_manual(values=ml.colors)

```


```{r}
ml.gg
```

Summarize by group, expanded
```{r}
options(dplyr.summarise.inform=FALSE)
n.folds <- 10
ml.exp.atime <- atime::atime(
  N=as.integer(10^seq(2, 7, by=0.5)),
  setup={
    loss.dt <- data.table(
      name="loss", 
      fold=rep(1:n.folds, each=2*N),
      loss=rnorm(2*N*n.folds),
      set=rep(c("subtrain","validation"),each=N),
      epoch=1:N)
    key.dt <- data.table(loss.dt, key=c("set","epoch","fold"))
  },
  seconds.limit=seconds.limit,
  "[.data.table(no key)"={
    loss.dt[, .(
      loss_length=.N,
      loss_mean=mean(loss),
      loss_sd=sd(loss)
    ), by=.(set, epoch)]
  },
  "[.data.table(key)"={
    key.dt[, .(
      loss_length=.N,
      loss_mean=mean(loss),
      loss_sd=sd(loss)
    ), by=.(set, epoch)]
  },
  "stats::aggregate"={
    res <- stats::aggregate(
      loss ~ set + epoch, 
      loss.dt, 
      function(values)list(c(
        loss_length=length(values),
        loss_mean=mean(values), 
        loss_sd=sd(values))))
    data.frame(
      subset(res, select=-loss), 
      do.call(rbind, res$loss))
  },
  "dplyr::summarise"={
    loss.dt |> 
      dplyr::group_by(set, epoch) |> 
      dplyr::summarise(
        loss_length=length(loss),
        loss_mean=mean(loss), 
        loss_sd=sd(loss))
  },
  "collapse::fsummarise"={
    loss.dt |> 
      collapse::fgroup_by(set, epoch) |> 
      collapse::fsummarise(
        loss_length=length(loss),
        loss_mean=mean(loss), 
        loss_sd=sd(loss))
  })

```
```{r}
ml.exp.refs <- atime::references_best(ml.exp.atime)
ml.exp.pred <- predict(ml.exp.refs)
ml.exp.colors <- c(
  "collapse::fsummarise"="#5AAE61",
  "dplyr::summarise"="#9970AB",
  "[.data.table(key)"="#D6604D",
  "[.data.table(no key)"="#B6604D",
  "stats::aggregate"="deepskyblue")
ml.exp.gg <- plot(ml.exp.pred)+
  theme(text=element_text(size=20))+
  ggtitle(sprintf("Mean,SD,Length over %d real numbers, N times", n.folds))+
  scale_x_log10("N = number of Mean,SD,Length to compute")+
  scale_y_log10("Computation time (seconds)
median line, min/max band
over 10 timings")+
  facet_null()+
  scale_fill_manual(values=ml.exp.colors)+
  scale_color_manual(values=ml.exp.colors)
```

```{r}
ml.exp.gg
```

